Memory-Efficient Mixed-Precision Implementations\\for Robust Explicit Model Predictive Control

### 2019-04-12 emsoft-2019/Cont_Syn.tex
section 4

If the region error is smaller than $\Delta$, we still have an error budget \texttt{errorAct} left
over to implement the control actions in finite precision (line 11).
For the precision assignment, we first call Daisy with the expressions for the
control actions given by \texttt{F} and \texttt{G}, and the remaining error
budget as the error bound (line 12).
Then we call Daisy with the expressions of the region bounds given by \texttt{H}
and \texttt{K}, now with the error bound $\varepsilon_Q$ (line 13).
In both cases, Daisy determines a quantization of the input matrices (Daisy also
determines the fixed-point formats of the arithmetic operations, which we ignore here).
Daisy either computes a minimal uniform fixed-point precision (word length),
i.e. it searches for the smallest uniform precision which still satisfies the
error bound, or it returns a mixed-precision assignment, where each value and
intermediate operation can potentially have different word length.
In our experiments in~\autoref{sec:experiments}, we compare the two modes and show
that mixed-precision leads to smaller memory footprints.

@@

\paragraph{Implementation}

>>>>

If the region error is smaller than $\Delta$, we still have an error budget \texttt{errorAct} left
over to implement the control actions in finite precision (line 11).
For the precision assignment, we first call Daisy with the expressions for the
control actions given by \texttt{F} and \texttt{G}, and the remaining error
budget as the error bound (line 12).
Then we call Daisy with the expressions of the region bounds given by \texttt{H}
and \texttt{K}, now with the error bound $\varepsilon_Q$ (line 13).

In both cases, Daisy determines a quantization of the input matrices (Daisy also
determines the fixed-point formats of the arithmetic operations, which we ignore here).
Daisy either computes a minimal uniform fixed-point precision (i.e. word length),
or it returns a mixed-precision assignment, where each value and
intermediate operation can potentially have a different word length.
To determine the minimal uniform precision, Daisy performs a linear search.
It starts from the smallest precision (1 bit) and checks whether it satisfies the
error bound. If not, the precision is increased by 1 bit and the first precision
that satisfies the error bound is returned.
Minimal uniform precision is used as a starting point for mixed-precision tuning,
which attempts to assign a lower precision to some variables. 
In our experiments in~\autoref{sec:experiments}, we compare the two modes and show
that mixed precision leads to smaller memory footprints.

@@

\subsection{Implementation}

### 2019-04-12 emsoft-2019/Error_Analysis.tex
fix
section 4

We first present our error analysis assuming given fixed-point word lengths,
before explaining our optimization algorithm which determines suitable word
lengths fully automatically in~\autoref{sec:Controller_Synthesis}.
\eva{The `p' below is the fixed/uniform word length? If so, please explain here}
\eva{Should we mention here the paper which does the LP-based analysis, i.e.
contrast our approach?}

In the same way described in \cite{imperialrmpc}, the output of the controller can be affected by two main sources errors which an error
analysis has to capture: 
\begin{itemize}
  \item[(i)] \emph{incorrect region selection}: an incorrect control action can be chosen
    due to analog-digital conversion and quantization of $H_i$ and $K_i$,

@@

We thus have to ensure that these errors remain below the error bound $\delta$
used to design the robust MPC controller:
\begin{align}\label{eq:delta}
  \max_{\forall i,j. \mathit{neighbour}(\mathcal{R}_{i},\mathcal{R}_{j})}|u_{i}-u_{j}| + \max_{\forall i.\mathcal{R}_{i}\in\mathcal{R}} err(u_{i})_{p} \le \delta
\end{align}
where $i,j$ are the indexes of two generic regions $\mathcal{R}_{i}$ and
$\mathcal{R}_{j}$, \\ $\mathit{neighbour}(\mathcal{R}_{i},\mathcal{R}_{j})$ states that
two regions share at least a point, $u_{i}$ is the
control action for controller $i$, and $err(u_{i})_{p}$ is
the approximation error for computing the activation function $u_{i}$ in fixed precision using p bits format.
\paragraph{Incorrect Region Selection}

The first part of the error in~\autoref{eq:delta} captures the error due to
selection of the wrong control action, i.e. with infinite-precision arithmetic
and in the absence of any uncertainty from hardware measurements we would choose
control action $u_i$, but due to uncertainties we actually choose control action $u_j$:
\begin{align} \label{eq:maximization}
  &\max_{\forall i,j.\mathit{neighbour}(\mathcal{R}_{i},\mathcal{R}_{j})}|u_{i}-u_{j}|= \\
  &\max_{\forall i,j.\mathit{neighbour}(\mathcal{R}_{i},\mathcal{R}_{j})}|F_{i}\statevar+G_{i} - (F_{j}\statevar+G_{j})|\nonumber
\end{align}

@@

A naive computation of this error would consider all possible values of $x_k$.
This would lead to a gross over-approximation, however, since the controller
will choose a wrong region only in a relative narrow neighbourhood, or
\emph{tube}, around the boundary between two regions $R_i$ and $R_j$.
In~\autoref{eq:maximization}, we thus constrain $x_k$ to be within an appropriate tube:
\begin{equation}\label{eq:tube}
(\mathit{border}_{i,j} \ge -\varepsilon) \land (\mathit{border}_{i,j} \le \varepsilon)
\end{equation}
where $\mathit{border}_{i,j}$ constrains $x_k$ to be on the border between two regions
$R_i$ and $R_j$ and $\varepsilon$ captures the uncertainty in computing the
correct region.

\eva{The below, I believe, is somehow defining $border_{i,j}$, but I don't understand
how. Please explicitly define $border_{i,j}$.}
In formula $\exists\; m,n \;$such that:
\begin{equation}
\mathit{border}_{i,j}=(H_{i}\statevar-K_{j})_{m} = (H_{j}\statevar-K_{j})_{n}
\end{equation}
where $m$ and $n$ are the indexes of the matching border respectively in $\mathcal{R}_{i}$ and $\mathcal{R}_{j}$.

The width of the tube $\varepsilon$, i.e. the uncertainty in computing $x_k$ and
thus the correct region, is due to two sources of errors: analog-to-digital
conversion $\varepsilon_{A/D}$ and the quantization of region bounds
$\mathcal{R}$ in memory $\varepsilon_{Q}$:
\begin{equation}\label{eq:epsilontot}
  \varepsilon=\varepsilon_{A/D}+\varepsilon_{Q}
\end{equation}

Analog conversion happens just before the controller receives the sensor input from the plant
and is given by:
\begin{equation*}
\varepsilon_{A/D}=\frac{V_{cc}}{2^{p}-1}
\end{equation*}
where $V_{cc}$ is the reference voltage of the converter (e.g. typically 5V), p is the number of bits available to quantize the analog signal, and $2^{p}$ is the resolution of the converter.

\eva{which bits, who determines these?}

While $\varepsilon_{A/D}$ is intrinsic to the capabilities of the device,
$\varepsilon_{Q}$ depends on the precision used to store the boundaries, and it is user defined at compile time. This
second error can be tuned based on a trade-off between accuracy of the storage and
memory storage required~\cite{memoryMPC}.
%
$\varepsilon_{Q}$ is given by:
\begin{equation}\label{eq:quantizationlines}
  \varepsilon_{Q} > |(H-\hat{H}_{p})\qstatevar+(K-\hat{K}_{p})|
\end{equation}
\eva{Shouldn't this equation have some absolute values in it?}
\eva{The equation above does not give any hint for how to compute it. I would prefer to add the inequality to \autoref{eq:epsilontot} and give an equality here that explicitly shows how to compute this. Then also rephrase the below}

The inequality guarantees that the distance between any hyperplane in infinite precision (H and K), and its counter-part quantized in p bits precision ($\hat{H}_{p}$ and $\hat{K}_{p}$) is bounded by $\varepsilon_{Q}$.

\paragraph{Approximate Control Output}

Computing $u_{i}$ itself introduces imprecision because $u_i$ needs to be evaluated 
in finite precision arithmetic:
\begin{equation}\label{eq:fperror}
  err(u_{i})_{p}=|(\hat{F}_{p}-F)\statevar+(\hat{G}_{p}-G)|
\end{equation}
where $\hat{F}$ and $\hat{G}$ represent the rounded values (in p bits) for the infinite precision values $F$ and $G$.

In (\ref{eq:fperror}) the domain for \statevarmath are all the points in region $i$, together with all the values in the tube surrounding region $i$. 
In this way we compute the error also for those points, that belong to a neighbor of $i$, but because of $\varepsilon$ they are erroneously located inside region $i$.

\eva{This section so far says nothing about how we actually compute these errors (incl. with which tools). 
This should come here}
\paragraph{Implementation}

The designer of the controller fixes the value for $epsilon$ at compile time before starting the analysis. $\varepsilon_{Q}$ is also a user dependent parameter fixed at compile time. The value for $\varepsilon_{Q}$ depends on the total amount of memory we can dedicate to the storage of H and K on the target controller.

The first part of equation (7) is solved in MATLAB. Since the function $u_{i}-u_{j}$ is linear, and we define the tube as a convex region, we need to evaluate the function only at the corner points of the tube and keep the value with the maximal magnitude.

The second part of equation (7), and the precision required for hyperplanes storage are both computed in Daisy. 
The way we use the analyzer is the following: we give an error bound in input to Daisy together with the coefficients we want to quantize (H, K or F, G). Daisy iteratively tries different uniform precisions for the quantization of the coefficients, and stops when it finds the minimal word-length such that the error in input is satisfied.
Daisy starts the search from unary word-length (1bit), and increases the precision with a one by one step at each iteration. It stops when the current (uniform) configuration introduces an approximation error that is bounded by the input error. This minimal word-length is going to be the baseline for the mixed-precision tuning phase.

Our analysis returns the total memory (in bits) required by the configuration detected in Daisy. The designer can then tune the epsilon value based on the memory consumption of the current configuration.

>>>>

We first present our error analysis assuming given fixed-point word lengths,
before explaining our optimization algorithm which determines suitable word
lengths fully automatically in~\autoref{sec:Controller_Synthesis}.
While we do not employ the error analysis independently of controller and
precision synthesis, we note that it is an alternative way of bounding errors to
existing methods~\cite{imperialrmpc}.

For simplicity of notation, we assume a uniform word length $p$ in this section,
though our algorithm (and implementation) support mixed-precision as well.

The output of an explicit MPC controller can be affected by two main sources
errors, which an error analysis has to capture\cite{imperialrmpc}:
\begin{itemize}
  \item[(i)] \emph{incorrect region selection}: 
    due to analog-digital conversion and quantization of $H_i$ and $K_i$,
    the controller can choose an incorrect control action $u_j$ instead of $u_i$

@@

We have to ensure that these errors remain below the error bound $\Delta$
used to design the robust MPC controller:
\begin{align}\label{eq:delta}
  \max_{\forall i,j. \mathit{neighbor}(\mathcal{R}_{i},\mathcal{R}_{j})}|u_{i}-u_{j}| + \max_{\forall i.\mathcal{R}_{i}\in\mathcal{R}} err(u_{i})_{p} \le \Delta
\end{align}
where $i,j$ are the indexes of two distinct regions $\mathcal{R}_{i}$ and
$\mathcal{R}_{j}$, \\ $\mathit{neighbor}(\mathcal{R}_{i},\mathcal{R}_{j})$
states that two regions share at least a point, $u_{i}$ is the control action
for controller $i$, and $err(u_{i})_{p}$ captures the approximation error for
computing the control action $u_{i}$ in fixed precision $p$.

\subsection{Incorrect Region Selection}

The first part of the error in~\autoref{eq:delta} captures the error due to
selecting of the wrong control action. That is, with infinite-precision arithmetic
and in the absence of any uncertainty from hardware measurements we would choose
control action $u_i$, but due to uncertainties we actually choose control action $u_j$:
\begin{align} \label{eq:maximization}
  \max_{\forall i,j.\mathit{neighbor}(\mathcal{R}_{i},\mathcal{R}_{j})}|u_{i}-u_{j}|&= \\
  \max_{\forall i,j.\mathit{neighbor}(\mathcal{R}_{i},\mathcal{R}_{j})}&|F_{i}\statevar+G_{i} - (F_{j}\statevar+G_{j})|\nonumber
\end{align}

@@

A naive computation of this error would consider all possible values of $x_k$.
This would lead to a gross over-approximation, however, since the controller
will choose a wrong region only in a relative narrow neighborhood, or
\emph{tube}, around the boundary between two regions $R_i$ and $R_j$.

The border between two regions $\mathcal{R}_{i}$ and $\mathcal{R}_{j}$ is defined by 
their corresponding boundaries:
\begin{equation}
\mathit{border}_{i,j} :=(H_i \statevar - K_i ) - (H_j \statevar - K_j)
\end{equation}
for appropriate matching borders $H_i, K_i$ and $H_j, K_j$.

In~\autoref{eq:maximization}, we thus constrain $\statevar$ to be within an appropriate tube:
\begin{equation}\label{eq:tube}
|\mathit{border}_{i,j}| \le \varepsilon
\end{equation}
where $\varepsilon$ captures the uncertainty in computing the
correct region and thus the width of the tube.
%
The width of the tube $\varepsilon$ is due to two sources of errors:
analog-to-digital conversion $\varepsilon_{A/D}$ and the quantization of region
bounds in memory $\varepsilon_{Q}$:
\begin{equation}\label{eq:epsilontot}
  \varepsilon=\varepsilon_{A/D}+\varepsilon_{Q}
\end{equation}

Analog conversion happens just before the controller receives the sensor input from the plant
and is given by:
\begin{equation*}
\varepsilon_{A/D}=\frac{V_{cc}}{2^{p}-1}
\end{equation*}
where $V_{cc}$ is the reference voltage of the converter (e.g. typically 5V), $p$
is the number of bits available to quantize the analog signal, and $2^{p}$ is
the resolution of the converter.

While $\varepsilon_{A/D}$ is intrinsic to the capabilities of the device,
$\varepsilon_{Q}$ depends on the precision used to store the boundaries:
\begin{equation}\label{eq:quantizationlines}
  \varepsilon_{Q} = |(H_i - \hat{H}_i) \qstatevar + (K_i - \hat{K}_i)|
\end{equation}
\eva{For which index is this evaluated, $i$, $j$ or both?}
That is, $\varepsilon_{Q}$ bounds the distance between any hyperplane in infinite
precision (H and K), and its counter-part quantized in $p$ bits precision
($\hat{H}_{p}$ and $\hat{K}_{p}$).
This second error can be tuned providing a trade-off between accuracy and
memory storage required.

\subsection{Approximate Control Output}

Once a control action $u_i$ is chosen, computing $u_{i}$ itself introduces
imprecision because $u_i$ needs to be evaluated in finite-precision arithmetic:
\begin{equation}\label{eq:fperror}
  err(u_{i})_{p} = | (F_i - \hat{F}_{i}) \statevar + ( G_i - \hat{G}_{i} ) |
\end{equation}
where $\hat{F}$ and $\hat{G}$ represent the quantized values (in $p$ bits) for
the infinite precision values $F$ and $G$.

\autoref{eq:fperror} is evaluated for all \statevarmath that are inside the
region $\mathcal{R}_i$, together with all the values in the tube surrounding
region $\mathcal{R}_i$.
In this way we compute the error also for those points that belong to a neighbor
of $\mathcal{R}_i$, but because of finite-precision errors they are erroneously
mapped to control action $u_i$.

\subsection{Implementation}

The incorrect region error (\autoref{eq:maximization}) is computed using
Matlab's multi-parametric toolbox. Due to the linearity of function
$u_{i}-u_{j}$, and because we defined the tube as a convex region, it suffices
to evaluate the function only at the corner points of the tube, and keep the
result with the maximal magnitude.
\eva{This needs more details. Which function(s) and algorithm from Matlab are used?}
\eva{How do you determine which regions are neighbors? resp. which borders are matching?}

$\varepsilon_Q$ and $err(u_{i})_{p}$ are computed by Daisy. For this we encode
the expressions in~\autoref{eq:quantizationlines} and \autoref{eq:fperror}
respectively as straight-line arithmetic expressions and specify the constraints
on the domain of $\statevar$ in the precondition.
Daisy performs a dataflow analysis to determine the finite-precision roundoff
errors. This error computation differs from the one employed by
\citet{imperialrmpc} which phrases the roundoff error computation as a (relaxed)
mixed integer programming problem.

While our actual synthesis algorithm (\autoref{sec:Controller_Synthesis}) does
not compute the errors $\varepsilon_Q$ and $err(u_{i})_{p}$ explicitly, the
error verification is used internally by Daisy to determine a suitable precision
$p$ and can also be called explicitly without the precision optimization.

% The approximation error and the precision tuning for both borders and active
% functions are computed in Daisy. $\varepsilon_{Q}$ is given in input to the
% analysis. After we compute the incorrect region in MATLAB we obtain the value
% for max error $err(u_{i})$ from (8). Then we give the errors and the vectors F,G
% and H,K in input to Daisy. We used Daisy as a verification tool: we provide in
% input the expression we want to quantize (e.g. $\mathcal{R}_{i}$) together with
% the error bound we can tolerate (e.g. $\varepsilon_{Q}$), and Daisy returns a
% precision configuration (uniform or mixed) such that the input error is
% satisfied. The way it finds the right configuration for uniform precision is the
% following: it starts by checking whether the input expression, quantized in
% unary uniform precision (1bit), introduces an error bounded by the input error.
% If this is not the case, Daisy increases the precision format by 1 bit and check
% whether the error now is satisfied. This unary increment runs inside a loop that
% terminates when the first valid format is matched. This uniform word-length is
% going to be the baseline for the mixed-precision tuning phase. The
% mixed-precision algorithm keeps a list of variables to quantize, and collects
% different candidate configurations that satisfy the input error. It returns the
% best configuration based on the evaluation of a quality function.
